---
title: "final project 553"
output: html_document
runtime: shiny
---




initial research question: how does amount of agricultural land and/or rural land, and percent of urban population impact the number of refugees in countries in 2018?

loading in needed libraries
```{r}
library(tidyverse)
library(shiny)
#install.packages("ggpubr")
library(ggpubr)
```


loading in the data  note -- these datasets are available in my github, and are originally sourced from the world data bank. I did do some editing in excel before use
```{r}
lit_rt <- read_csv("adult_lit_rate_2018.csv")
agri_prc <- read_csv("agri_land_percent_2018.csv")
life_expec <- read_csv("life_expectancy_at_birth_2018.csv")
ref_pop <- read_csv("refugee_population_by_country_of_asylum_2018.csv")
rural_pop_prc <- read_csv("rural_pop_total_percent_2018.csv")
tot_pop <- read_csv("total_population_2018.csv")
urbn_pop_prc <- read_csv("urban_pop_total_2018.csv")
slums_prc_urbn <- read_csv("urban_slums_pop_2018.csv")

```

combining the data into a usable form
```{r}
#inner join to keep the data tidy
a <- inner_join(lit_rt, agri_prc, by=c("Country Code", "Country Name"))
b <- inner_join(a, life_expec, by=c("Country Code", "Country Name"))
c <- inner_join(b, ref_pop, by=c("Country Code", "Country Name"))
d <- inner_join(c, rural_pop_prc, by=c("Country Code", "Country Name"))
e <- inner_join(d, tot_pop, by=c("Country Code", "Country Name"))
f <- inner_join(e, urbn_pop_prc, by=c("Country Code", "Country Name"))
df <- inner_join(f, slums_prc_urbn, by=c("Country Code", "Country Name"))
df

#getting the refugee population as a percentage
df <- df %>% mutate(`Refugee Perc` = `refugee population country asylum`/`Total Population`) %>% select(-c(`refugee population country asylum`, `Percent Urban Pop in Slums`, `Total Population`, `Agricultural Land Perc`, `Country Code`))
df


#renaming it to proper variable names
df <- df %>% rename(country=`Country Name`)%>%rename(lit_rate = `Literacy Rate 2018`)%>%rename(life_exp=`Life Expectancy Years`)%>%rename(rural_pop=`Rural Population Percent`)%>%rename(urbn_pop=`Urban Pop Percent`)%>%rename(ref_perc=`Refugee Perc`)


```

basic eda to begin to address the question
the variable of interest is Refugee Percent
```{r}
#there look to be three outliers, lets see what those are 
df[order(-df$ref_perc),]
#apparently, the countries/territories with the highest asylum seekers/refugees is West Bank and Gaza, Jordan, and Lebanon. Upon some minor internet research, it would appear that most of the refugees in these places are Palestinian refugees, and further that the literacy rate in Palestine is 97.5%
#making a new dataset excluding these three, just to see what the graphs would look like
h <- df%>%filter(ref_perc<0.1)
h[order(-h$ref_perc),]
```

```{r}
#plots
#refugee percent by adult literacy
ggplot(df, aes(ref_perc, lit_rate))+geom_point()+geom_smooth(se=FALSE, method='lm')
ggplot(h, aes(ref_perc, lit_rate))+geom_point()+geom_smooth(se=FALSE, method='lm')
#refugee percent by life expectancy
ggplot(df, aes(ref_perc, life_exp))+geom_point()+geom_smooth(se=FALSE, method='lm')
ggplot(h, aes(ref_perc, life_exp))+geom_point()+geom_smooth(se=FALSE, method='lm')
#refugee percent by rural population
ggplot(df, aes(ref_perc, rural_pop))+geom_point()+geom_smooth(se=FALSE, method='lm')
ggplot(h, aes(ref_perc, rural_pop))+geom_point()+geom_smooth(se=FALSE, method='lm')
#refugee percent by urban population
ggplot(df, aes(ref_perc, urbn_pop))+geom_point()+geom_smooth(se=FALSE, method='lm')
ggplot(h, aes(ref_perc, urbn_pop))+geom_point()+geom_smooth(se=FALSE, method='lm')
```

I need two categorical variables, so I will make abv_lit, which is 1 for if the country has above average adult literacy and 0 if not, and abv_life with the same for average life expectancy
```{r}
#get the average of literacy and life expectancy
j <- df%>%select(-country)
l <- colMeans(j, na.rm = TRUE)
#using ifelse to create the dummy vars
#literacy first
k <- ifelse(df$lit_rate>l[1], 1, 0)
#next life expectancy
m <- ifelse(df$life_exp>l[2], 1, 0)
#now I join these columns to the main dataset and rename them
df <- cbind(df, k)%>%cbind(m)%>%rename(abv_lit=k)%>%rename(abv_life=m)
#and make a smaller and more manageable dataset
df_1 <- df%>%select(country, rural_pop, urbn_pop, ref_perc, abv_lit, abv_life)
```

second set of eda
```{r}
ggplot(df_1, aes(rural_pop, ref_perc))+geom_point()+geom_smooth(se=FALSE, method='lm')
ggplot(df_1, aes(urbn_pop, ref_perc))+geom_point()+geom_smooth(se=FALSE, method='lm')

ggplot(df_1)+geom_boxplot(aes(factor(abv_lit), ref_perc))
ggplot(df_1)+geom_boxplot(aes(factor(abv_life), ref_perc))

```

beginning to work on the shiny app
```{r}
ui <- fluidPage(
  #title of the whole site
  titlePanel("Refugee Data"),
  #adding a sidebar layout
  sidebarLayout(
    #adding the actual sidebar
    sidebarPanel(
      textInput(inputId="name", "What's your name?"),
  textOutput(outputId = "greeting"),
  br(),
  textInput(inputId = "title", "What should the title of this project be?"),
  textOutput("title1"),
  br(),
  verbatimTextOutput("hover_text_u"),
  verbatimTextOutput("hover_text_r")
      #first input -- Which other variable to compare to refugee percent
      #radioButtons(inputId="cont_var", "Variable for Comparison", c("Population Percentage Urban" = "urbn_pop", "Population Percentage Rural" = "rural_pop")),
      #second input -- How many of the top inputs to remove
      #radioButtons(inputId = "perc_rem", "What top percent of refugee population countries to remove:", c("None"=1, "1 percent or more"=0.01, "2.5 percent or more"=0.025, "5 percent or more"=0.05)),
      #third input -- Which categorical variable to use
      #radioButtons(inputId = "cat_var", "Categorical Variable to Use", c("Above average adult literacy"="abv_lit", "Above average life expectancy"="abv_life"))
    ),
    mainPanel(
      tabsetPanel(type = "tabs",
                  tabPanel("Regression Plot for Urban Population", plotOutput(outputId = "reg_plot_u", hover = hoverOpts(id="plot_hover_u"))),
                  tabPanel("Regression Equation for Urban Population", verbatimTextOutput(outputId = "reg_eq_u")),
                  tabPanel("Regression Plot for Rural Population", plotOutput(outputId = "reg_plot_r", hover = hoverOpts(id="plot_hover_r"))),
                  tabPanel("Regression Equation of Rural Population", verbatimTextOutput(outputId = "reg_eq_r")),
                  tabPanel("Box-Plot of Life Expectancy", plotOutput(outputId = "box_plot")),
                  #tabPanel("Removed", tableOutput(outputId = "removed_rows")),
                  tabPanel("Inferential Statistics", verbatimTextOutput(outputId = "inf_stat"))
      )
    )
  )
)

server <- function(input, output, session){
  output$reg_plot_u <- renderPlot({
    ggplot(df, aes(x=urbn_pop, y=ref_perc))+geom_point()+geom_smooth(method="lm", se=FALSE)+xlab("Percent of Population in Urban Areas")+ylab("Percent of Population Made Up of Refugees")+stat_regline_equation()+labs(subtitle = "We fail to reject the hypothesis that there is no relationship between these two variables")
  })
  output$reg_plot_r <- renderPlot({
    ggplot(df, aes(x=rural_pop, y=ref_perc))+geom_point()+geom_smooth(method="lm", se=FALSE)+xlab("Percent of Populaiton in Rural Areas")+ylab("Percent of Population Made Up of Refugees")+stat_regline_equation()+labs(subtitle = "We fail to reject the hypothesis that there is no relationship between these two variables")
  })
  output$reg_eq_u <- renderPrint({
    reg_u <- lm(df$ref_perc ~ df$urbn_pop)
    summary(reg_u)
  })
  output$reg_eq_r <- renderPrint({
    reg_r <- lm(df$ref_perc ~ df$rural_pop)
    summary(reg_r)
  })
  output$box_plot <- renderPlot({
    ggplot(df)+geom_boxplot(aes(factor(abv_life), ref_perc), outlier.shape = NA)+scale_y_continuous(limits = c(0,0.025))+xlab("Above Average Life Expectancy (1=above, 0=below)")+ylab("Percent of Population Made Up of Refugees")
  })
  output$inf_stat <- renderPrint({
    print("T test for comparing if refugee populations are similar between population life expectancies. The two groups are NOT significantly different.")
    t.test(df$ref_perc~df$abv_life, mu=0, alt="two.sided", var.eq=F)
  })
  output$greeting <- renderText({
    paste0("Hello ", input$name, "!")
  })
  output$title1 <- renderText({
    paste0("The title of this project is: ", input$title)
  })
  output$hover_text_u <- renderPrint({
    if(!is.null(input$plot_hover_u)){
            hover=input$plot_hover_u
            paste0(df$urbn_pop, ",", df$ref_perc)
        }
  })
  output$hover_text_r <- renderPrint({
    if(!is.null(input$plot_hover_r)){
            hover=input$plot_hover_r
            paste0(df$rural_pop, ",", df$ref_perc)
        }
  })
  
}

#server <- function(input, output, session){
 # output$reg_plot <- renderPlot({
  #  df%>%ggplot(aes(x=input$cont_var, y=ref_perc))+geom_point()
#  })
#}


shinyApp(ui=ui, server=server)


```












